### 2020-09-30 PNAS_PostReview/Manuscript_PostReview.Rmd
PNAS post-review resubmission formatting and packaging

---
bibliography: manuscript.bib
csl: pnas.csl
site: "bookdown::bookdown_site"
output:
  bookdown::pdf_document2:
    toc: FALSE
indent: TRUE
monofont: "Times New Roman"
fontsize: 12pt
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \doublespacing
---

@@

Predator specialization has often been considered an evolutionary ‘dead-end’ due to the constraints associated with the evolution of morphological and functional optimizations throughout the organism. However, in some predators, these changes are localized in separate structures dedicated to prey capture. One of the most extreme cases of this modularity can be observed in siphonophores, a clade of pelagic colonial cnidarians that use tentilla (tentacle side branches armed with nematocysts) exclusively for prey capture. Here we study how siphonophore specialists and generalists evolve, and what morphological changes are associated with these transitions. To answer these questions, we: (1) measured 29 morphological characters of tentacles from 45 siphonophore species, (2) mapped these data to a phylogenetic tree, and (3) analyzed the evolutionary associations between morphological characters and prey type data from the literature. Instead of a dead-end, we found that siphonophore specialists can evolve into generalists, and that specialists on one prey type have directly evolved into specialists on other prey types. Our results show that siphonophore tentillum morphology has strong evolutionary associations with prey type, and suggest that shifts between prey types are linked to shifts in the morphology, mode of evolution, and genetic correlations of tentilla and their nematocysts. The evolutionary history of siphonophore specialization helps build a broader perspective on predatory niche diversification via morphological innovation and evolution. These findings contribute to understanding how specialization and morphological evolution have shaped present-day food webs.

@@

Predatory specialization is often associated with the evolution of modifications in the morphology of the prey capture apparatus. Specialization has been considered an evolutionary ‘dead-end’ due to the constraints associated with these morphological changes. However, in predators like siphonophores, armed with modular structures used exclusively for prey capture, this assumption is challenged. Our results show that siphonophores can evolve generalism and new prey-type specializations by modifying the morphological states, modes of evolution, and genetic correlations between the parts of their prey capture apparatus. These findings demonstrate how studying open-ocean non-bilaterian predators can reveal novel patterns and mechanisms in the evolution of specialization. Understanding these evolutionary processes is fundamental to the study of food-web structure and complexity.

@@

  Siphonophores (Cnidaria: Hydrozoa) are a clade of gelatinous, colonial organisms that swim in the open ocean, feeding on a wide diversity of prey (often fish, crustaceans, and jellyfish). Siphonophores carry modular structures that are exclusively used for prey capture: the tentilla (Fig. \@ref(figure1)). The tentilla have great morphological variation across species [@mapstone2014global]. Together with their well understood function, this makes them an ideal system to study the relationships between functional traits and prey specialization. Like a head of coral, a siphonophore is a colony bearing many feeding polyps (Fig. \@ref(figure1)). Each feeding polyp has a single tentacle, which branches into a series of tentilla (side branches). Like other cnidarians, siphonophores capture prey with nematocysts, harpoon-like stinging capsules borne within specialized cells known as cnidocytes. Unlike the prey capture apparatus of most other cnidarians, siphonophore tentacles carry their cnidocytes in extremely complex and organized batteries [@skaer1988formation] which are located in their tentilla. While nematocyst batteries and clusters in other cnidarians are simple static scaffolds for cnidocytes, siphonophore tentilla have their own reaction mechanism, triggered upon encounter with prey. When it fires, a tentillum undergoes an extremely fast conformational change that wraps it around the prey, maximizing the surface area of contact for nematocysts to fire on the prey [@Mackie:1987uy]. In addition, some species have elaborate fluorescent and bioluminescent lures on their tentilla to attract prey with aggressive mimicry [@purcell1980influence;@haddock2005bioluminescent;@haddock2015fluorescent].

@@

![\label{figure1} Siphonophore anatomy. A - *Nanomia* sp. siphonophore colony (photo by Catriona Munro). B, C - Illustration of a *Nanomia* colony, gastrozooid, and tentacle closeup (by Freya Goetz). D - *Nanomia* sp. Tentillum illustration and main parts. E - Differential interference contrast micrograph of the tentillum illustrated in D. F - Nematocyst types (illustration reproduced with permission from Mapstone 2014), hypothesized homologies, and locations in the tentillum. Undischarged to the left, discharged to the right.](Figures_final/figure1.jpg)

@@

  Specialization has been thought to be an evolutionary ‘dead-end’, meaning that specialized lineages are unlikely to evolve into generalists or to shift the resource for which they are specialized [@futuyma1988evolution;@siddall1993phylogeny;@hougen1994interactions; @robinson1996trade;@kelley1998specialization]. However, recent studies have found that interspecific competition can favor the evolution of generalists from specialists [@stireman2005evolution;@johnson2009competition;@nosil2002transition] and specialist resource switching [@hoberg2008macroevolutionary;@hardy2014specialization]. In addition to studying relationships with morphology, we seek to identify what evolutionary transitions in trophic niche breadth are prevalent in these open-ocean tactile predators. To do so, we examine three alternative scenarios of siphonophore trophic specialization: (1) predatory specialists evolved from generalist ancestors; (2) predatory specialists evolved from specialist ancestors which targeted different resources, switching their primary prey type; and (3) predatory generalists evolved from specialist ancestors. These scenarios are non-exclusive, and each could apply to different transitions along the siphonophore phylogeny.

@@

  In this study, we present an extensive morphological characterization of tentilla and their nematocysts across a broad variety of shallow and deep-sea siphonophore species using modern imaging technologies, summarize the literature on siphonophore diets, expand the phylogenetic tree of siphonophores by combining ribosomal gene sequences from a broad range of taxa with a transcriptome-based backbone tree, and explore the evolutionary histories and correlations between diet, tentillum, and nematocyst characters. Our results suggest that siphonophores can evolve new specializations and generalism by modifying the phenotypes and genetic correlations in their prey capture apparatus. These findings show how studying elusive non-bilaterian predators can challenge traditional views on the evolution of predatory specialization.

@@

![\label{figure4} Bayesian time-tree inferred from 18S + 16S concatenated sequences and constrained to be congruent with a published transcriptome phylogeny. Branch lengths estimated using a relaxed molecular clock. Species names in red indicate replicated representation in the morphology data. All data were publicly available, apart from new sequences produced for *Thermopalia taraxaca* and *Frillaglma vityazi* (bold). Nodes labeled with Bayesian posteriors (BP). Green circles indicate BP = 1. Blue circles indicate nodes constrained to be congruent with Munro *et al.* (2018). Tips with black squares indicate the species with transcriptomes used in Munro *et al.* (2018). Tips with purple squares indicate genus-level correspondence to taxa included in Munro *et al.* (2018). The main clades are labeled: with black bars for described taxonomic units, and grey bars for operational phylogenetic designations.](Figures_final/figure4.png){height=600px}

@@

![\label{figure5} Left - Subset phylogeny showing the mapped feeding guild regimes that were used to inform the *OUwie* analyses. Right - Grid showing the prey items consumed from which the feeding guild categories were derived. Diet data were obtained from the literature review, available in the Dryad repository.](Figures_final/figure5.jpg)

@@

```{r tab1, echo=FALSE, out.width='100%'}
knitr::include_graphics('Final_tables/table1.png')
```
\

@@

```{r tab2, echo=FALSE, out.width='60%'}
knitr::include_graphics('Final_tables/table2.png')
```

@@

Several studies [@futuyma1988evolution;@siddall1993phylogeny;@hougen1994interactions; @robinson1996trade;@kelley1998specialization] have suggested that resource specialization can be an irreversible dead-end due to the constraints posed by extreme phenotypic specialization. Our results show that this is not the case for siphonophores, where the prey type on which they specialize has shifted at least 5 times. We find no support for any transitions from generalist to specialist (scenario 1, as described in the Introduction). We do find support for at least 3 instances of specialists switching from one prey type to another prey type, (scenario 2) and two switches from specialist to generalist (scenario 3). This is consistent with the findings of recent studies on phytophagous insects [@nosil2002transition], where the rate of evolution from generalists to specialists is comparable to the reverse, thus specialization does not limit further evolution. Our results are also consistent with analyses of lepidopterans [@hardy2014specialization], where specialized resource switching is the primary transition type while niche breadth remains fairly constant. The evolutionary history of tentilla shows that siphonophores are an example of trophic niche diversification via morphological innovation and evolution, which allowed transitions between specialized trophic niches. In more familiar predators, the prey capture apparatus (such as claws and jaws) is well integrated in the body, leading to trade-offs and whole-body adaptations for feeding specialization. The extreme modularity of the siphonophore prey capture apparatus could release them from constraints typically imposed by adaptation to ecological specialization. This evolutionary mechanism is particularly important in a deep open-ocean ecosystem, which is a relatively homogeneous physical environment, where the primary niche heterogeneity available is the potential interactions between organisms [@robison2004deep].

@@

Our results unambiguously show that tentillum morphology evolved with diet and strongly support deviations from the generalist-to-specialist evolution scenario. However, the conclusions we can draw from these analyses are limited in several ways. The biggest challenge at present is the sparse dietary data available in the literature. Additional dietary data could reveal transitions from generalists to specialists we were unable to detect for two reasons: some of the taxa in our dataset have a very limited number of feeding observations, which could lead to apparent specialization; and some of the taxa not included in our dataset could be undiscovered generalists. When interpreting these results, it is also important to remember that diet is also dependent on environmental prey availability. In addition, selectivity differences across siphonophore species could be also driven by other phenotypes not accounted for in this study. Finally, further observations on behavior, digestion biochemistry, and toxin composition are necessary to assess their relative importance in determining diet.

@@

Most studies on the evolution of predation have focused on vertebrate systems with an integrated feeding apparatus serving multiple functions. This has led to a narrow understanding of the evolutionary outcomes of specialization, where extreme morphological evolution constrains further shifts in their ecology. Siphonophores differ in many ways from commonly-known predators, using modular weapons for prey capture (the tentilla) that are fully decoupled from other structures and body functions. Our analysis of the evolutionary history of dietary specialization and morphological change in these elusive animals has revealed notable deviations from traditional expectations. While much of the feeding ecology literature focuses on how predatory generalists evolve into predatory specialists, in siphonophores we find predatory specialists can evolve into generalists, and that specialists on one prey type have directly evolved into specialists on other prey types. We find that the character states, evolutionary optima, and genetic correlations of many morphological characters have evolved following these ecological shifts. We find that the relationships between form and ecology hold across a large set of siphonophore taxa and characters. These findings are central to understanding the evolutionary mechanisms driving the emergence of food web complexity.

@@

*Statistical analyses* – We used a series of phylogenetic comparative methods to test the evolutionary hypotheses presented in this study. We reconstructed ancestral states using ML (R phytools::anc.ML [@revell2012phytools]), and stochastic character mapping (R phytools::make.simmap) for categorical characters. For more details on the data wrangling prior to these analyses, please see the Supplementary Methods. R scripts available in the Dryad repository.

@@

Data will be available from the Dryad Digital Repository: [http://dx.doi.org/10.5061/dryad.NNNN](http://dx.doi.org/10.5061/dryad.[NNNN])
Supplementary Materials (Supplementary Methods and figures) are appended following this manuscript text, data available in [https://github.com/dunnlab/tentilla_morph/tree/master/Supplementary_materials](https://github.com/dunnlab/tentilla_morph/tree/master/Supplementary_materials)

@@

This work was supported by the Society of Systematic Biologists (Graduate Student Award to A.D.S.); the Yale Institute of Biospheric Studies (Doctoral Pilot Grant to A.D.S.); and the National Science Foundation (Waterman Award to C.W.D., and NSF-OCE 1829835 to C.W.D., S.H.D.H., and C. Anela Choy). A.D.S. was supported by a Fulbright Spain Graduate Studies Scholarship. We wish to thank the crew and scientists of the R/V Western Flyer, who participated in the collection of many of the specimens used in this study. We also want to thank Lynne Christianson and Shannon Johnson from the Monterey Bay Aquarium Research Institute for their assistance in the field as well as for sequencing some of the species included in this phylogeny. In addition, we wish to thank Lourdes Rojas, Daniel Drew, and Eric Lazo-Wasem for their assistance in imaging the fixed specimens and managing the collections. We thank Dennis Pilarczyk for organizing the prey selectivity data, Michael Landis for helping design the Bayesian analyses, and Joaquin Nunez for reviewing this manuscript. Furthermore, we thank Elizabeth D. Hetherington and C. Anela Choy for collating the data on siphonophore feeding and for reviewing the manuscript. Finally, we thank Philip Pugh, who confirmed many of our specimen identifications and taught us valuable knowledge about siphonophores.

>>>>

---
bibliography: manuscript.bib
csl: pnas.csl
site: "bookdown::bookdown_site"
output:
  bookdown::word_document2:
    toc: FALSE
indent: TRUE
monofont: "Times New Roman"
fontsize: 12pt
header-includes:
  - \usepackage{setspace}
  - \doublespacing
---

@@

Predator specialization has often been considered an evolutionary ‘dead-end’ due to the constraints associated with the evolution of morphological and functional optimizations throughout the organism. However, in some predators, these changes are localized in separate structures dedicated to prey capture. One of the most extreme cases of this modularity can be observed in siphonophores, a clade of pelagic colonial cnidarians that use tentilla (tentacle side branches armed with nematocysts) exclusively for prey capture. Here we study how siphonophore specialists and generalists evolve, and what morphological changes are associated with these transitions. To answer these questions, we: (1) measured 29 morphological characters of tentacles from 45 siphonophore species, (2) mapped these data to a phylogenetic tree, and (3) analyzed the evolutionary associations between morphological characters and prey type data from the literature. Instead of a dead-end, we found that siphonophore specialists can evolve into generalists, and that specialists on one prey type have directly evolved into specialists on other prey types. Our results show that siphonophore tentillum morphology has strong evolutionary associations with prey type, and suggest that shifts between prey types are linked to shifts in the morphology, mode of evolution, and evolutionary correlations of tentilla and their nematocysts. The evolutionary history of siphonophore specialization helps build a broader perspective on predatory niche diversification via morphological innovation and evolution. These findings contribute to understanding how specialization and morphological evolution have shaped present-day food webs.

@@

Predatory specialization is often associated with the evolution of modifications in the morphology of the prey capture apparatus. Specialization has been considered an evolutionary ‘dead-end’ due to the constraints associated with these morphological changes. However, in predators like siphonophores, armed with modular structures used exclusively for prey capture, this assumption is challenged. Our results show that siphonophores can evolve generalism and new prey-type specializations by modifying the morphological states, modes of evolution, and evolutionary correlations between the parts of their prey capture apparatus. These findings demonstrate how studying open-ocean non-bilaterian predators can reveal novel patterns and mechanisms in the evolution of specialization. Understanding these evolutionary processes is fundamental to the study of food-web structure and complexity.

@@

  Siphonophores (Cnidaria: Hydrozoa) are a clade of gelatinous, colonial organisms that swim in the open ocean, feeding on a wide diversity of prey (often fish, crustaceans, and jellyfish). Siphonophore colonies have a modular body plan with different zooids specialized for different tasks. This modularity extends also within the feeding gastrozooids, which carry modular structures on the tentacle that are exclusively used for prey capture: the tentilla (Fig. \@ref(figure1)). The tentilla have great morphological variation across species [@mapstone2014global]. Together with their well understood function, this makes them an ideal system to study the relationships between functional traits and prey specialization. Like a head of coral, a siphonophore is a colony bearing many feeding polyps (Fig. \@ref(figure1)). Each feeding polyp has a single tentacle, which branches into a series of tentilla (side branches). Like other cnidarians, siphonophores capture prey with nematocysts, harpoon-like stinging capsules borne within specialized cells known as cnidocytes. Unlike the prey capture apparatus of most other cnidarians, siphonophore tentacles carry their cnidocytes in extremely complex and organized batteries [@skaer1988formation] which are located in their tentilla. While nematocyst batteries and clusters in other cnidarians are simple static scaffolds for cnidocytes, siphonophore tentilla have their own reaction mechanism, triggered upon encounter with prey. When it fires, a tentillum undergoes an extremely fast conformational change that wraps it around the prey, maximizing the surface area of contact for nematocysts to fire on the prey [@Mackie:1987uy]. In addition, some species have elaborate fluorescent or bioluminescent lures on their tentilla to attract prey with aggressive mimicry [@purcell1980influence;@haddock2005bioluminescent;@haddock2015fluorescent].

@@

![\label{figure1} Siphonophore anatomy. A - *Nanomia* sp. siphonophore colony (photo by Catriona Munro). B, C - Illustration of a *Nanomia* colony, gastrozooid, and tentacle closeup (by Freya Goetz). D - *Nanomia* sp. Tentillum illustration and main parts. E - Differential interference contrast micrograph of the tentillum illustrated in D. F - Nematocyst types (illustration reproduced with permission from Mapstone 2014), hypothesized homologies, and locations in the tentillum. Undischarged to the left, discharged to the right.](~/tentilla_morph/Figures_final/figure1.jpg)

@@

  Specialization in resource acquisition and use has often been presented as an evolutionary ‘dead-end’ [@siddall1993phylogeny;@hougen1994interactions; @robinson1996trade;@kelley1998specialization; summarized by @futuyma1988evolution]. The concept of a ‘dead-end’ can be problematic because it conflates very different macroevolutionary patterns. These patterns can pertain to the clade, such as higher extinction rates or lower diversification rates, or to the evolutionary lock-in of lineages to particular attributes. Here we exclusively focus on this last sense, in which feeding specialization can be considered a ‘dead-end’ if lineages that are feeding specialists do not give rise to feeding generalists or specialists on other prey. However, recent studies have found that ecological mechanisms such as interspecific competition can favor the evolution of generalists from specialists [@stireman2005evolution;@johnson2009competition;@nosil2002transition] and specialist resource switching [@hoberg2008macroevolutionary;@hardy2014specialization]. In addition to studying relationships with morphology, we seek to identify what evolutionary transitions in trophic niche breadth are prevalent in these open-ocean tactile predators. To do so, we examine three alternative scenarios of siphonophore trophic specialization: (1) predatory specialists evolved from generalist ancestors; (2) predatory specialists evolved from specialist ancestors which targeted different resources, switching their primary prey type; and (3) predatory generalists evolved from specialist ancestors. These scenarios are non-exclusive, and each could apply to different transitions along the siphonophore phylogeny.

@@

  In this study, we present an extensive morphological characterization of tentilla and their nematocysts across a broad variety of shallow and deep-sea siphonophore species using modern imaging technologies, summarize the literature on siphonophore diets, expand the phylogenetic tree of siphonophores by combining ribosomal gene sequences from a broad range of taxa with a transcriptome-based backbone tree, and explore the evolutionary histories and correlations between diet, tentillum, and nematocyst characters. Our results suggest that siphonophores can evolve new specializations and generalism by modifying the phenotypes and evolutionary correlations in their prey capture apparatus. These findings show how studying elusive non-bilaterian predators can challenge traditional views on the evolution of predatory specialization.

@@

![\label{figure4} Bayesian time-tree inferred from 18S + 16S concatenated sequences and constrained to be congruent with a published transcriptome phylogeny. Branch lengths estimated using a relaxed molecular clock. Species names in red indicate replicated representation in the morphology data. All data were publicly available, apart from new sequences produced for *Thermopalia taraxaca* and *Frillaglma vityazi* (bold). Nodes labeled with Bayesian posteriors (BP). Green circles indicate BP = 1. Blue circles indicate nodes constrained to be congruent with Munro *et al.* (2018). Tips with black squares indicate the species with transcriptomes used in Munro *et al.* (2018). Tips with purple squares indicate genus-level correspondence to taxa included in Munro *et al.* (2018). The main clades are labeled: with black bars for described taxonomic units, and grey bars for operational phylogenetic designations.](~/tentilla_morph/Figures_final/figure4.png){height=600px}

@@

![\label{figure5} Left - Subset phylogeny showing the mapped feeding guild regimes that were used to inform the *OUwie* analyses. Right - Grid showing the prey items consumed from which the feeding guild categories were derived. Diet data were obtained from the literature review, available in the Dryad repository.](~/tentilla_morph/Figures_final/figure5.jpg)

@@

```{r tab1, echo=FALSE, out.width='100%'}
knitr::include_graphics('~/tentilla_morph/Final_tables/table1.png')
```
\

@@

```{r tab2, echo=FALSE, out.width='60%'}
knitr::include_graphics('~/tentilla_morph/Final_tables/table2.png')
```

@@

Several studies [@futuyma1988evolution;@siddall1993phylogeny;@hougen1994interactions; @robinson1996trade;@kelley1998specialization] have suggested that resource specialization can be an irreversible state due to the constraints posed by extreme phenotypic specialization. Our results show that this is not the case for siphonophores, where the prey type on which they specialize has shifted at least 5 times. We find no support for any transitions from generalist to specialist (scenario 1, as described in the Introduction). We do find support for at least 3 instances of specialists switching from one prey type to another prey type, (scenario 2) and two switches from specialist to generalist (scenario 3). This is consistent with the findings of recent studies on phytophagous insects [@nosil2002transition], where the rate of evolution from generalists to specialists is comparable to the reverse, thus specialization does not limit further evolution. Our results are also consistent with analyses of lepidopterans [@hardy2014specialization], where specialized resource switching is the primary transition type while niche breadth remains fairly constant. These results show how an ancestor specialized for feeding on a particular prey can still give rise to multiple lineages and to novel feeding guilds, including generalists. For example, we find that, Eucladophora, a large clade of siphonophores that contains the majority of extant species and diverse feeding strategies, arose from a most recent common ancestor that was a specialist on large crustacean prey. Though ‘evolutionary dead-end’ is a problematic term that could apply to several evolutionary patterns, this result is inconsistent with multiple specific uses of this term.

The evolutionary history of tentilla shows that siphonophores are an example of trophic niche diversification via morphological innovation and evolution, which allowed transitions between specialized trophic niches. In more familiar predators, the prey capture apparatus (such as claws and jaws) is well integrated in the body, leading to trade-offs and whole-body adaptations for feeding specialization. The extreme modularity of the siphonophore prey capture apparatus could release them from constraints typically imposed by adaptation to ecological specialization. This evolutionary mechanism is particularly important in a deep open-ocean ecosystem, which is a relatively homogeneous physical environment, where the primary niche heterogeneity available is the potential interactions between organisms [@robison2004deep].

@@

Our results unambiguously show that tentillum morphology evolved with diet and strongly support deviations from the generalist-to-specialist evolution scenario. However, the conclusions we can draw from these analyses are limited in several ways. The biggest challenge at present is the sparse dietary data available in the literature. Additional dietary data could reveal transitions from generalists to specialists we were unable to detect for two reasons: some of the taxa in our dataset have a very limited number of feeding observations, which could lead to apparent specialization; and some of the taxa not included in our dataset could be undiscovered generalists. When interpreting these results, it is important to remember that diet is also dependent on environmental prey availability, which was not available in most of the sources we used (except Purcell 1981 & Purcell 1984). We integrated published dietary data collected using different methods bearing different inherent biases. Gut content inspections (used in the majority of our literature sources) are very effective at detecting small hard-bodied prey, but can fail to detect rapidly-digested soft-bodied prey. ROV observations can be biased towards large (often gelatinous) prey, and can easily oversee smaller prey items. In addition, selectivity differences across siphonophore species could be also driven by other phenotypes not accounted for in this study. Finally, further observations on behavior, digestion biochemistry, and toxin composition are necessary to assess their relative importance in determining diet.

We hypothesize that siphonophores are able to evolve from specialization for one prey type to other states due to their prey-capture apparatus being extremely independent from the rest of the bodies in terms of location and function. We also hypothesize that the homogeneous midwater environment they live in favors the evolution of extreme morphological adaptations for prey capture. Testing these hypotheses will be interesting in its own right, and also give a better sense of how generalizable our results are beyond siphonophores. It is important to note that our hypotheses only apply to organisms with access to a broad-enough diversity of resources (such as prey, or hosts) on which to specialize, and only when said resources pose distinct challenges that require anatomical modifications. Otherwise, there would not be enough variation to detect these patterns in the first place.

@@

Most studies on the evolution of predation have focused on vertebrate systems with an integrated feeding apparatus serving multiple functions. This has led to a narrow understanding of the evolutionary outcomes of specialization, where extreme morphological evolution constrains further shifts in their ecology. Siphonophores differ in many ways from commonly-known predators, using modular weapons for prey capture (the tentilla) that are fully decoupled from other structures and body functions. Our analysis of the evolutionary history of dietary specialization and morphological change in these elusive animals has revealed notable deviations from traditional expectations. While much of the feeding ecology literature focuses on how predatory generalists evolve into predatory specialists, in siphonophores we find predatory specialists can evolve into generalists, and that specialists on one prey type have directly evolved into specialists on other prey types. We find that the character states, evolutionary optima, and evolutionary correlations of many morphological characters have evolved following these ecological shifts. We find that the relationships between form and ecology hold across a large set of siphonophore taxa and characters. These findings are central to understanding the evolutionary mechanisms driving the emergence of food web complexity.

@@

*Statistical analyses* – We used a series of phylogenetic comparative methods to test the evolutionary hypotheses presented in this study. We reconstructed ancestral states using ML (R phytools::anc.ML [@revell2012phytools]), and stochastic character mapping (R phytools::make.simmap) for categorical characters. When reconstructing the evolutionary history of feeding guilds, we fitted our SIMMAP model under the agnostic assumption that a generalist diet poses morphological challenges that are as distinct as each specialization is from each other. Thus, we do not impose any *a priori* constraints or weighting in the model for what state transition rates are permissible, letting the data determine the parameters. For more details on the data wrangling prior to these analyses, please see the Supplementary Methods. R scripts available in the Dryad repository.

@@

Data are available from the Dryad Digital Repository: [http://dx.doi.org/10.5061/dryad.p2ngf1vp2](http://dx.doi.org/10.5061/dryad.p2ngf1vp2)
Supplementary Materials (Supplementary Methods and figures) are appended following this manuscript text, data available in [https://github.com/dunnlab/tentilla_morph/tree/master/Supplementary_materials](https://github.com/dunnlab/tentilla_morph/tree/master/Supplementary_materials)

@@

This work was supported by the Society of Systematic Biologists (Graduate Student Award to A.D.S.); the Yale Institute of Biospheric Studies (Doctoral Pilot Grant to A.D.S.); and the National Science Foundation (Waterman Award to C.W.D., and NSF-OCE 1829835 to C.W.D., OCE-1829805 to S.H.D.H., and OCE-1829812 to C. Anela Choy). A.D.S. was supported by a Fulbright Spain Graduate Studies Scholarship. We wish to thank the crew and scientists of the R/V Western Flyer, who participated in the collection of many of the specimens used in this study. We also want to thank Lynne Christianson and Shannon Johnson from the Monterey Bay Aquarium Research Institute for their assistance in the field as well as for sequencing some of the species included in this phylogeny. In addition, we wish to thank Lourdes Rojas, Daniel Drew, and Eric Lazo-Wasem for their assistance in imaging the fixed specimens and managing the collections. We thank Dennis Pilarczyk for organizing the prey selectivity data, Michael Landis for helping design the Bayesian analyses, and Joaquin Nunez for reviewing this manuscript. Furthermore, we thank Elizabeth D. Hetherington and C. Anela Choy for collating the data on siphonophore feeding and for reviewing the manuscript. Finally, we thank Philip Pugh, who confirmed many of our specimen identifications and taught us valuable knowledge about siphonophores.
